---
title: "tables-graphs"
format: html
---

```{r}
library(tidyverse)
library(here)
Estimated_tariff_impact <- read_csv(here("data","Estimated_Impact_Trump_Tariffs.csv"))
Revenue_effects <- read_csv(here("data", "Revenue_Effects_Trump_Tariffs.csv"))
```

```{r}
est_tariff_impact <- Estimated_tariff_impact |>
  mutate(tariff = ...1) |>
  select(tariff, GDP, `Capital Stock`, `Pre-Tax Wages`,`Hours Worked Converted to Full-Time Equivalent Jobs`) |>
  filter(!is.na(tariff)) |>
  mutate(GDP = str_sub(GDP,3,5)) |>
  mutate(GDP = str_replace(GDP, "ss", ".05")) |>
  mutate(GDP = as.numeric(GDP)) |>
  mutate(`Capital Stock` = str_sub(`Capital Stock`,3,5)) |>
  mutate(`Capital Stock` = str_replace(`Capital Stock`,"ss",".05")) |>
  mutate(`Capital Stock` = as.numeric(`Capital Stock`)) |>
  mutate("GDP_dec" = GDP) |>
  mutate("Capital_stock_dec" = `Capital Stock`) |>
  select(tariff, GDP_dec, Capital_stock_dec)
head(est_tariff_impact)
```

Capital Stock: Capital stock refers to the total number of shares that a company is authorized to issue, as outlined in its corporate charter. It represents the maximum amount of shares the company can ever have outstanding (That can be held by investors). 

```{r}
Revenue_effects <- Revenue_effects |>
  mutate(tariff = `Conventional Revenue`) |>
  select(-`Conventional Revenue`) |>
  relocate(tariff, .before = `2025`) |>
  filter(!is.na(tariff)) |>
  filter(tariff != "Dynamic Revenue")
```

What countries GDP has/will been most effected by the tariffs

```{r}
Rev_effects_long <-Revenue_effects |>
  pivot_longer(cols = c(2:12), names_to = "year", values_to = "est. revenue effect (in billions)") |>
  mutate(`est. revenue effect (in billions)` = str_sub(`est. revenue effect (in billions)`, 2,10)) |>
  mutate(`est. revenue effect (in billions)` = as.numeric(`est. revenue effect (in billions)`)) |>
  filter(year != "2025-2034") |>
  mutate(year = as.numeric(year))
```

Visualization of Revenue Effects

```{r}
Rev_effects_long |>
  ggplot(aes(x=year, y =`est. revenue effect (in billions)`, color = tariff)) +
  geom_line() +
  labs(x = "Year", y = "Estimated Revenue effect (in billions)", title = "Estimated Revenue Effects of Each Tariff 2025-2034")
```

Consider here creating another rev effects graph
```{r}
Rev_effects_long |>
  filter(tariff == "IEEPA Fentanyl/\"Reciprocal\" China") %>%
  filter(`est. revenue effect (in billions)`==!is.na)
  ggplot(aes(x=year, y =`est. revenue effect (in billions)`)) +
  geom_line(color = "orange") +
  labs(x = "Year", y = "Estimated Revenue effect (in billions)", title = "Estimated Revenue Effects of Each Tariff 2025-2034")
```


Visualization of GDP decrease due to different tariffs

```{r}
ggplot(est_tariff_impact, aes(x = GDP_dec, fill = tariff)) +
  geom_histogram() +
  labs(x = "Percent of Estimated GDP Decrease", title = "Estimated GDP Impact of Each Tariff")
```

Visualization of Capital Stock decrease of different tariffs

```{r}
ggplot(est_tariff_impact, aes(x = Capital_stock_dec, fill = tariff)) +
  geom_histogram() +
  labs(x = "Percent of Estimated Capital Stock Decrease", title = "Estimated Capital Stock Impact of Each Tariff")
```

Correlation of GDP and capital stock decrease of different tariffs 

```{r}
ggplot(est_tariff_impact, aes(x = GDP_dec, y = Capital_stock_dec, color = tariff)) +
  geom_point() +
  labs(x = "Percent GDP decrease", y = "Percent Capital Stock dec.")

# Figure out how to make the points bigger if at all possible
# Try and reason which tariffs are missing and why that is (if it matters) explain why or why not
```

